---
title: "EDA_Project"
author: "Dusty P"
date: "May 31, 2018"
output: html_document
---

```{r echo=FALSE, message=FALSE, warning=FALSE, setup}
knitr::opts_knit$set(root.dir = normalizePath("C:/Users/Dusty/Documents/coding/projects/Udacity/Data Analysis/eda/EDA_Project"))

# load the ggplot graphics package and the others
library(ggplot2)
library(GGally)
library(scales)
library(memisc)
library(gridExtra)
library(RColorBrewer)
library(bitops)
library(RCurl)

cuberoot_trans = function() trans_new('cuberoot', transform = function(x) x^(1/3),
                                      inverse = function(x) x^3)
```

# Exploration of White Wines by Dustin Pianalto

This report explores a dataset containing chemical information and ratings on almost 4900 white wine tastings.

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
wqw <- read.csv('wineQualityWhites.csv')
# because the first column is just row numbers I am going to remove it
wqw <- subset(wqw, select = -X)
```

# Univariate Plots Section

```{r echo=FALSE, Data_Dimensions}
dim(wqw)
```

```{r echo=FALSE, Data_Structure}
str(wqw)
```

```{r echo=False, Data_Summary}
summary(wqw)
```

Our Data consists of 11 numerical variables and one Integer attribute which is the output with almost 4900 observations

```{r echo=FALSE, quality_histogram}
ggplot(aes(x = quality), data = wqw) + 
  geom_histogram(binwidth = 1)
```

The distribution of the quality seems fairly normal with a peak at 6

```{r echo=FALSE, alcohol_histogram}
ggplot(aes(x = alcohol), data = wqw) + 
  geom_histogram(binwidth = .1)
```

The Alcohol seems to be slightly long tailed, I want to see what it is like with a log transformation.

```{r echo=FALSE, alcohol_histogram}
ggplot(aes(x = alcohol), data = wqw) + 
  geom_histogram(binwidth = .005) +
  scale_x_log10()
```

```{r echo=FALSE, fixed.acidity_histogram}
ggplot(aes(x = fixed.acidity), data = wqw) + 
  geom_histogram(binwidth = .1)
```

The fixed.acidity definately has some outliers but besides that has a pretty normal distribution.

```{r echo=FALSE, fixed.acidity_summary}
summary(wqw$fixed.acidity)
```

Most Wines have a acidity between 6.3 and 7.3
I am going to plot the data again removing both the high and low 1% of values to remove the outliers.

```{r echo=FALSE, fixed.acidity_histogram}
ggplot(aes(x = fixed.acidity), data = wqw) + 
  geom_histogram(binwidth = .1) +
  xlim(quantile(wqw$fixed.acidity, 0.01), quantile(wqw$fixed.acidity, 0.99))
```

And we see a fairly normal distribution with a peak around 6.8 which matches both the median (6.8) and mean (6.855) from the summary above.

```{r echo=FALSE, volatile.acidity_histogram}
ggplot(aes(x = volatile.acidity), data = wqw) + 
  geom_histogram(binwidth = .01)
```

We have another long tailed distribution. I am going to plot again with a log_10 transformation this time.

```{r echo=FALSE, volatile.acidity_histogram}
ggplot(aes(x = volatile.acidity), data = wqw) + 
  geom_histogram(binwidth = .04) +
  scale_x_log10()
```

```{r echo=FALSE, citric.acid_histogram}
ggplot(aes(x = citric.acid), data = wqw) + 
  geom_histogram(binwidth = .01) +
  xlim(quantile(wqw$citric.acid, 0.01), quantile(wqw$citric.acid, 0.99))
```

There is an odd spike at about 0.49 I might want to look into that more later.

```{r echo=FALSE, residual.sugar_histogram}
ggplot(aes(x = residual.sugar), data = wqw) + 
  geom_histogram(binwidth = .1) +
  xlim(quantile(wqw$residual.sugar, 0.01), quantile(wqw$residual.sugar, 0.99))
```

Even with the top and bottom 1% removed the plot is still very long tailed

```{r echo=FALSE, residual.sugar_histogram}
p1 <- ggplot(aes(x = residual.sugar), data = wqw) + 
  geom_histogram(binwidth = .05) +
  scale_x_log10()
p2 <- ggplot(aes(x = residual.sugar), data = wqw) + 
  geom_histogram(binwidth = .01) +
  scale_x_log10(breaks = seq(0, 20, 2))
grid.arrange(p1, p2)
```

```{r echo=FALSE, residual.sugar_summary}
summary(wqw$residual.sugar)
```

Using a log_10 transform with a bin width of .05 indicates a bimodal distribution but if you decrease the binwidth to 0.01 it shows that while there are a lot of observations between ~4 and ~20 they are a lot more spread out and there are more individual of each value from ~0.5 to ~2. And we can see in the summary of the data that the median is 5.2 and the mean is 6.4 which puts both of them inbetween the two peaks.

```{r echo=FALSE, chlorides_histogram}
ggplot(aes(x = chlorides), data = wqw) + 
  geom_histogram(binwidth = .001) +
  xlim(0, quantile(wqw$chlorides, 0.97))
```

Here I just removed the top 3% of values to remove the long tail.

```{r echo=FALSE, sulfur.dioxide_histograms}
p1 <- ggplot(aes(x = free.sulfur.dioxide), data = wqw) + 
  geom_histogram(binwidth = 1) +
  xlim(0, quantile(wqw$free.sulfur.dioxide, 0.99))
p2 <- ggplot(aes(x = total.sulfur.dioxide), data = wqw) + 
  geom_histogram(binwidth = 1) +
  xlim(quantile(wqw$total.sulfur.dioxide, 0.01), quantile(wqw$total.sulfur.dioxide, 0.99))
grid.arrange(p1, p2)
```

I plotted the Free Sulphur Dioxide and Total Sulphur Dioxide together to save room and because they are related. Note the difference in scales on both axies.

```{r echo=FALSE, density_histogram}
ggplot(aes(x = density), data = wqw) + 
  geom_histogram(binwidth = .0001) +
  xlim(quantile(wqw$density, 0.01), quantile(wqw$density, 0.99))
```

```{r echo=FALSE, pH_histogram}
ggplot(aes(x = pH), data = wqw) + 
  geom_histogram(binwidth = .01)
```

The pH plot doesn't need any modification.

```{r echo=FALSE, sulphates_histogram}
ggplot(aes(x = sulphates), data = wqw) + 
  geom_histogram(binwidth = .01)
```



# Univariate Analysis

### What is the structure of your dataset?

There are 4898 samples in the dataset with 11 different variables and a resulting quality assesment. All of the variables are continuous number variables and the quality is an integer scale from 1 to 10 with max value of 9 and min of 3.

Observations:
 * The most common quality is 6 and it is a fairly normal distribution slightly skewed towards the low end.
 * Most of the variables are similar in distribution, most of them are long tailed but besides that have a fairly normal distribution.
 * There are a couple interesting features though, the Citric Acid has an odd spike around 4.9 and the Residual Sugar appears to be more of a bimodal distribution.
 
### What is/are the main feature(s) of interest in your dataset?

My main interest in this dataset is trying to determine which features have the greatest effect on the quality.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

I think that the Alcohol, Acidity, Density, and Ph will have the greatest impact on the quality.

### Did you create any new variables from existing variables in the dataset?

I did not create any new variables.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

I either log transformed or removed the outliers on most of the datapoints to better view the data as most of them were longtailed.

# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}
ggpairs(wqw, axisLabels = "none", diag = list(continuous = wrap("diagAxis", labelSize=2, gridLabelSize=1)), upper = list(continuous = wrap("cor", size = 1.8)), lower = list(continuous = wrap("smooth", alpha=0.2, color = "orange")))
```

Looking at this grid there are not very many variables that appear to correlate with each other which I find suprising as some of them seem like they should. I am going to explore some of them in more detail.

One pair of variables that look like they have some correlation are density and residual sugar so I am going to start with them.

```{r echo=FALSE, residual.sugar_vs_density}
ggplot(aes(x = residual.sugar, y = density), data = wqw) +
  geom_point(color = "orange")
```

Narrowing in on the main section and adding a smoothing line.

```{r echo=FALSE, residual.sugar_vs_density_mod}
ggplot(aes(x = residual.sugar, y = density), data = wqw) +
  geom_point(alpha=0.3, color = "orange") +
  xlim(0, 30) +
  ylim(0.987, 1.0025) +
  geom_smooth()
```

We can see a general trend as residual sugar increases the density also increases. Lets see both of these plotted against our output variable.

```{r echo=FALSE, quality_vs_density}
ggplot(aes(x = quality, y = density), data = wqw) +
  ylim(0.985, 1.005) +
  geom_point(alpha=0.1, position = position_jitter(h = 0), color = "blue") +
  geom_line(stat = 'summary', fun.y = mean, color = "blue") +
  geom_line(stat = 'summary', fun.y = median) +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.1), color = 'red', linetype = 2) +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.9), color = 'red', linetype = 2)
```

```{r echo=FALSE, quality_vs_residual.sugar}
ggplot(aes(y = residual.sugar, x = quality), data = wqw) +
  geom_point(alpha=0.1, color = "blue") +
  xlim(0, 10) +
  ylim(0, 30)
```

There doesn't seem to be any direct corelation between these variables and the quality. Lets look at some others.

```{r echo=FALSE, quality_vs_alcohol}
ggplot(aes(x = quality, y = alcohol), data = wqw) +
  geom_point(alpha=0.1, color = "blue")
```

```{r echo=FALSE, quality_vs_alcohol}
ggplot(aes(x = quality, y = alcohol), data = wqw) +
  geom_point(alpha=0.1, position = position_jitter(h = 0), color = "blue") +
  geom_line(stat = 'summary', fun.y = mean, color = "blue") +
  geom_line(stat = 'summary', fun.y = median) +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.1), color = 'red', linetype = 2) +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.9), color = 'red', linetype = 2)
```

Adding jitter to the alcohol plot reveals that there could possibly be a corelation to quality but it is very weak.

```{r echo=FALSE, quality_vs_fixed.acidity}
ggplot(aes(x = quality, y = fixed.acidity), data = wqw) +
  geom_point(alpha=0.1, position = position_jitter(h = 0), color = "blue")
```

```{r echo=FALSE, quality_vs_chlorides}
ggplot(aes(x = quality, y = chlorides), data = wqw) +
  geom_point(alpha=0.1, position = position_jitter(h = 0), color = "blue") +
  ylim(0, 0.1) +
  geom_line(stat = 'summary', fun.y = mean, color = "blue") +
  geom_line(stat = 'summary', fun.y = median) +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.1), color = 'red', linetype = 2) +
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.9), color = 'red', linetype = 2)
```

```{r echo=FALSE, quality_vs_tsd}
ggplot(aes(x = quality, y = total.sulfur.dioxide), data = wqw) +
  geom_point(alpha=0.1, position = position_jitter(h = 0), color = "blue")
```

Looking at these other variables shows that there is little to no relationship to the quality individually I think this will change when we start combining variables in the Multivariate plots.

One other interesting corelation that I want to look at is density vs alcohol.

```{r echo=FALSE, alcohol_vs_density}
ggplot(aes(x = alcohol, y = density), data = wqw) +
  geom_point(alpha=0.1, color = "blue") +
  ylim(0.985, 1.005) +
  geom_smooth()
```

Interestingly it appears that as the aocohol content increases the density decreases, this is the inverse of the residual sugar vs density that we plotted earlier. This probably has something to do with the fact that sugar is what the alcohol is created from so it would follow that as the alcohol increases the sugar and thence the density would decrease.

We can see this more directly by plotting residual sugar against alcohol.

```{r echo=FALSE, alcohol_vs_residual_sugar}
ggplot(aes(x = alcohol, y = residual.sugar), data = wqw) +
  geom_point(alpha=0.1, color = "blue") +
  ylim(0, 30) +
  geom_smooth()
```

We can see that there is a seemingly exponential relationship between alcohol and residual sugar.

```{r echo=FALSE, alcohol_vs_chlorides}
ggplot(aes(x = alcohol, y = chlorides), data = wqw) +
  geom_point(alpha=0.1, color = "blue") +
  ylim(0, 0.1) +
  geom_smooth()
```

There does seem to be a slight corelation between alcohol and chlorides.

```{r echo=FALSE, alcohol_vs_ph}
ggplot(aes(x = pH, y = alcohol), data = wqw) +
  geom_point(alpha=0.1, color = "blue") +
  geom_smooth()
```

```{r echo=FALSE, alcohol_vs_fixed_acidity}
ggplot(aes(x = fixed.acidity, y = alcohol), data = wqw) +
  geom_point(alpha=0.1, color = "blue") +
  geom_smooth()
```

```{r echo=FALSE, alcohol_vs_volatile_acidity}
ggplot(aes(x = volatile.acidity, y = alcohol), data = wqw) +
  geom_point(alpha=0.1, color = "blue") +
  geom_smooth()
```

There does not seem to be any correlation between our other features of interest.


# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

I discovered some interesting relationships between density, residual sugar and alcohol. The other features appear to have very little corelation to each other or to the quality. The other relationships that I noted are the ones that were expected. For instance the pH has a mild corelation to the fixed acidity although I expected a higher corelation. Same with total sulfur dioxide and free sulfur dioxide.

It does seem like there is a mild corelation between the quality and alcohol as well as quality and density which are 2 of the features I noted in the previous section. There also might be a slight relationship between quality and chlorides.

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

One relationship that I found interesting is between alcohol and chlorides as well as between chlorides and quality. I wonder if this will show itself more in the multivariate exploration.

### What was the strongest relationship you found?

By far the strongest relationship I found was between density and residual sugar.

# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!